Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-07-04 PandaEvolution.tex
minor Section III-A

% \ifreview
% Comments to address in this section:
% \begin{enumerate}
% 	\item REVIEWER 2: The authors spent quite a few paragraphs for the detail
% 	design and implementation of the PanDA WMS, but introduced few background
% 	of the LHC applications. According to the paper, I understand the
% 	applications running on Titan are a large amount of small independent
% 	jobs. Thus, it seems less cost-effective to run these jobs on a
% 	world-leading supercomputer like Titan which has a fast and expensive
% 	interconnection. It will be more interesting if the authors can compare
% 	the solution on Titan (proposed in this paper) with the original
% 	distributed computing environment and HPECloud from Google mentioned in
% 	the related work section.
% 	\item REVIEWER 5: In Section 3.2, GPUs are not considered in this work,
% 	which would be a waste for using HPC computing. Could this work be
% 	extended to include GPUs or other heterogenous resources?
% \end{enumerate}
% \fi
The upcoming LHC Run 3 will require more resources than the Worldwide LHC
Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
300,000 cores at over 100 Grid sites, with a peak performance of 0.3
petaFLOPS\@. This capacity will be sufficient for the planned analysis and
data processing, but it will be insufficient for the Monte Carlo production
workflow and any extra activity. To alleviate these challenges, ATLAS is
expanding the current computing model to include additional resources such as
the opportunistic use of supercomputers.

Generally, supercomputers are designed to support parallel computation that
requires runtime communication. Jobs are executed across multiple cores, each
core calculating a small part of the problem and communicating with other
cores via MPI\@. Accordingly, supercomputers have large number of worker
nodes, connected through a high-speed, low-latency dedicated network. Each
worker node has multicore CPUs, usually augmented with Graphics Processing
Units (GPUs) or other specialized coprocessors.

PanDA WMS has been designed to support distributed Grid computing. Executing
ATLAS workloads or workflows involves concurrent and/or sequential runs of
possibly large amount of jobs, each requiring no or minimal parallelization
and no runtime communication. Thus, computing infrastructure like WLCG have
been designed to aggregate large amount of computing resources across
multiple sites. While each site may deploy MPI capabilities, usually these
are not used to perform distributed computations.

Currently, ATLAS workloads do not require fast interconnects or specilized
coprocessors but supercomputers tend not to reach 100\% utilization due to
the scheduling of jobs requiring large amount of resources. This offers the
possibility to execute ATLAS-like workloads on supercomputers to increase
utilization and reducing the waste of available resources.

We developed a single-point solution to better understand the problem space
of enabling a WMS designed for HTC to execute production workflows on
resources designed to support HPC\@. The PanDA team developed a job broker to
support the execution of part of the ATLAS production Monte Carlo workflow on
Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).

@@

Titan's architecture, configuration and policies poses several challenges to
the integration with PanDA\@. The default deployment model of PanDA Pilot is
unfeasible on Titan: PanDA Pilot is required to contact the Job Dispatcher of
the PanDA Server to pull jobs to execute, but this is not possible on Titan
because worker nodes do not offer outbound network connectivity. Further,
Titan does not support PanDA's security model based on certificates and
virtual organizations, making PanDA's approach to identity management also
unfeasible. While Titan's data transfer nodes (DTNs) offer wide area network
data transfer, an integration with ATLAS DDM is beyond the functional and
administrative scope of the current prototyping phase. Finally, the specific
characteristics of the execution environment, especially the absence of local
storage on the worker nodes and modules tailored to Compute Node Linux,
require re-engineering of ATLAS application frameworks.

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows
can be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
partitioned into independent jobs, Titan worker nodes can be used to execute
up to 16 concurrent payloads, one per each available core. Given these
constraints and challenges, the type of task most suitable for execution at
the moment on Titan is Monte Carlo detector simulation. This type of task is
mostly computational-intensive, requiring less than 2GB of RAM at runtime and
small input data. Detector simulation tasks in ATLAS are performed via
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 detector simulation toolkit~\cite{agostinelli2003geant4}. These tasks
account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
candidate for offloading.

Detector simulation is part of the ATLAS production Monte Carlo (MC)
workflow~\cite{rimoldi2006atlas}.%,de2013delphes,ritsch2014atlas}. 
The MC workflow consists of four main stages: event generation, detector
simulation, digitization, and reconstruction. Event generation creates sets
of particle four-momenta via different generators, e.g.,
PYTHIA,%~\cite{sjostrand2006pythia}, 
HERWIG,%~\cite{corcella2001herwig} 
and many others. Geant4 simulates the ATLAS detector and the interaction
between that detector and particles. Each interaction creates a so-called hit
and all hits are collected and passed on for digitalization, where hits are
further processed to mimic the readout of the detector. Finally,
reconstruction operates local pattern recognition, creating high-level
objects like particles and jets.

>>>>

% \ifreview
% Comments to address in this section:
% \begin{enumerate}
% 	\item REVIEWER 2: The authors spent quite a few paragraphs for the detail
% 	design and implementation of the PanDA WMS, but introduced few background
% 	of the LHC applications. According to the paper, I understand the
% 	applications running on Titan are a large amount of small independent
% 	jobs. Thus, it seems less cost-effective to run these jobs on a
% 	world-leading supercomputer like Titan which has a fast and expensive
% 	interconnection. It will be more interesting if the authors can compare
% 	the solution on Titan (proposed in this paper) with the original
% 	distributed computing environment and HPECloud from Google mentioned in
% 	the related work section.
% 	\item REVIEWER 5: In Section 3.2, GPUs are not considered in this work,
% 	which would be a waste for using HPC computing. Could this work be
% 	extended to include GPUs or other heterogenous resources?
% \end{enumerate}
% \fi
The upcoming LHC Run 3 will require more resources than the Worldwide LHC
Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
300,000 cores at more than 100 Grid sites, with a peak performance of 0.3
petaFLOPS\@. This capacity will be sufficient for the planned analysis and
data processing, but it will be insufficient for the Monte Carlo production
workflow and any extra activity. To alleviate these challenges, ATLAS is
expanding the current computing model to include additional resources such as
the opportunistic use of supercomputers.

Supercomputers are designed to support parallel computation that
requires runtime communication. Jobs are executed across multiple nodes, each
node calculating a small part of the problem and communicating with other
node, typically via MPI\@. Accordingly, supercomputers have large number of worker
nodes, connected through a high-speed, low-latency dedicated network. Each
worker node has multicore CPUs, usually augmented with Graphics Processing
Units (GPUs) or other specialized co-processors.

PanDA WMS has been designed to support distributed Grid computing. Executing
ATLAS workloads or workflows involves concurrent and/or sequential runs of
possibly large number of jobs, each requiring minimal, if any parallelization
and no runtime communication. Thus, computing infrastructure like WLCG have
been designed to aggregate large amount of computing resources across
multiple sites. While each site may deploy MPI capabilities, usually these
are not used to perform distributed computations.

Currently, ATLAS workloads do not require fast interconnects or specilized
co-processors, but supercomputers tend not to reach 100\% utilization due to
the scheduling of jobs requiring large amount of resources. This offers the
possibility to execute ATLAS-like workloads on supercomputers to increase
utilization and reducing the waste of available resources.

We developed a single-point solution to better understand the problem space
of enabling a WMS designed for HTC to execute production workflows on
resources designed to support HPC\@. The PanDA team developed a job broker to
support the execution of part of the ATLAS production Monte Carlo workflow on
Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
Computing Facility (OLCF).

@@

Titan's architecture, configuration and policies poses several challenges to
the deployment of PanDA\@. The default deployment model of PanDA Pilot is
unfeasible on Titan: PanDA Pilot is required to contact the Job Dispatcher of
the PanDA Server to pull jobs to execute, but this is not possible on Titan
because worker nodes do not offer outbound network connectivity. Further,
Titan does not support PanDA's security model based on certificates and
virtual organizations, making PanDA's approach to identity management 
unfeasible. While Titan's data transfer nodes (DTNs) offer wide area network
data transfer, an integration with ATLAS DDM is beyond the functional and
administrative scope of the current prototyping phase. Finally, the specific
characteristics of the execution environment, especially the absence of local
storage on the worker nodes and modules tailored to Compute Node Linux,
require re-engineering of ATLAS application frameworks.

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows
can be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
partitioned into independent jobs, Titan worker nodes can be used to execute
up to 16 concurrent payloads, one per each available core. Given these
constraints and challenges, the  Monte Carlo detector simulation task is most
suitable for execution on Titan at the moment. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and small
input data. Detector simulation tasks in ATLAS are performed via
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 detector simulation toolkit~\cite{agostinelli2003geant4}. These tasks
account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
candidate for offloading.

Detector simulation is part of the ATLAS production Monte Carlo (MC)
workflow~\cite{rimoldi2006atlas}. %,de2013delphes,ritsch2014atlas}. 
The MC workflow consists of four main stages: event generation, detector
simulation, digitization, and reconstruction. Event generation creates sets
of particle four-momenta via different generators, e.g.,
PYTHIA, %~\cite{sjostrand2006pythia}, 
HERWIG, %~\cite{corcella2001herwig} 
and many others. Geant4 simulates the ATLAS detector and the interaction
between the detector and particles. Each interaction creates a so-called hit
and all hits are collected and passed on for digitalization, where hits are
further processed to mimic the readout of the detector. Finally,
reconstruction operates local pattern recognition, creating high-level
objects like particles and jets.

### 2017-07-04 conclusion.tex
minor fixes to language, space saving etc.
minor Section III-A

Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 3: How extensible is this? Do you have a sense of backfill
	availability on other systems? Is this unique to Titan or Cray or more
	ubiquitous of an option?
\end{enumerate}

The deployment of PanDA Broker on Titan enabled distributed computing on a
leadership-class HPC machine at unprecedented scale. In the past 13 months,
PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of
the total number of detector events of the ATLAS production Monte Carlo
workflow. We described the implementation and execution process of PanDA WMS
(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}),
showing how they support and enable distributed computing at this scale on
Titan, a leadership-class HPC machine managed by OCLF at ORNL.

>>>>

\ifreview
Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 3: How extensible is this? Do you have a sense of backfill
	availability on other systems? Is this unique to Titan or Cray or more
	ubiquitous of an option?
\end{enumerate}
\fi

The deployment of PanDA Broker on Titan enabled distributed computing on a
leadership-class HPC machine at unprecedented scale. In the past 13 months,
PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of
the total number of detector events of the ATLAS production Monte Carlo
workflow. We described the implementation and execution process of PanDA WMS
(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}),
showing how they support and enable distributed computing at this scale on
Titan, a leadership-class HPC machine managed by OCLF.
